---
title: "Statistical Inference Course Project - Part 2"
author: "Ekaterina Abramova"
date: "21 January 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.
The following items should occur:  
        1. Load the ToothGrowth data and perform some basic exploratory data analyses  
        2. Provide a basic summary of the data.  
        3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.  
        4. State your conclusions and the assumptions needed for your conclusions.  

## 1.Load the ToothGrowth data and perform some basic exploratory data analyses
```{r}
library(datasets)
library(ggplot2)
```
Exploring the contents of the dataset
```{r}
head(ToothGrowth, 2)
str(ToothGrowth)
table(ToothGrowth$dose, ToothGrowth$supp)
```
## 2. Provide a basic summary of the data.  
The ToothGrowth data set consists of 60 observations of 3 variables:  
*len*: Tooth length (numeric variable)  
*supp*: Supplement type (factor variable with levels VC and OJ)  
*dose*: Dose in milligrams (numeric variable)  

We see a proportional characteristic: larger the dosage longer the tooth  
```{r echo = FALSE}
ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
    geom_bar(stat="identity") +
    facet_grid(. ~ supp) +
    xlab("Dosage in miligrams") +
    ylab("Tooth length")
```

## 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.  
```{r}
t.test(len ~ supp, data = ToothGrowth)
```
The p-value is greater than 0.05 and the confidence interval of the test contains zero so we can say that supplement types seems to have no impact on Tooth growth based on this test.  

We’ll compare tooth growth by dose
```{r}
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == .5, ])
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 1, ])
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 2, ])
```
##Conclusion:
Confidence testing shows that an increase in dosage is proportianal to longer tooth. But, with a p-value = 0.06 and having zero in the confidence interval we can not reject the null hypothesis. We can conclude that supplement delivery method has no effect on tooth length

