---
title: "Statistical Inference Course Project - Part 1"
author: "Ekaterina Abramova"
date: "21 January 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A simulation exercise
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials

### 1.Show the sample mean and compare it to the theoretical mean of the distribution.
We will create a simulation sample called simulated_samp each of size 40 drawn from exponential distribution with rate parameter $\lambda = 0.2$. 
Seed will be set at 333 to make the simulation reproducible.
```{r cars}
library(ggplot2)
# set seed for reproducability
set.seed(333)
# set lambda to 0.2
lambda <- 0.2
# 40 samples
n <- 40
# 1000 simulations
n.sim <- 1000
# simulate
simulated_samp <- matrix(rexp(n*n.sim, rate = lambda), nrow = n.sim, ncol = n)
# calculate mean of exponentials
simulated_samp_mean <- apply(simulated_samp, 1, mean)
mean(simulated_samp_mean)
```

```{r}
# analytical mean
theory_mean <- 1/lambda
theory_mean
```
The sample mean is 5.0453, which is close to the theoretical mean,  µ=1/$\lambda$=5

### 2.Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

```{r}
var(simulated_samp_mean)
```

```{r}
theory_var <- (1 / lambda ^ 2) / n
theory_var
```
The sample variance is 0.626, which is also close the theoretical variance 0.625



### 3.Show that the distribution is approximately normal
```{r}
data <- as.data.frame(simulated_samp_mean)
ggplot(data, aes(x = simulated_samp_mean)) + 
  geom_histogram(binwidth = 0.5, color = 'darkgreen', fill = 'white', aes(y = ..density..)) +
  stat_function( fun = dnorm, color = 'red', 
                args = list(mean = 5, sd = sqrt(0.626))) +
  xlab('Sample mean') +
  ylab('Density') +
  ggtitle('Comparison of the sample distribution and the theoretical distribution')+
  geom_vline(xintercept = c(mean(simulated_samp_mean), 5), size = c(1,1), 
               colour = c("darkgreen", "red"))

```
The above figure shows the distribution of the sample mean. It is approximately normal. The red density curve corresponds to N(5,0.626) density.
